### 2020-03-10 draft_analysis.Rmd
first commit

---
title: "COVID19 - Forecast analysis"
author: "PG"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
###import italian dataset updated 10 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
dat_csv$t<-1:dim(dat_csv)[1]
```

## The COVID dataset

The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.
We used a SEIR model to estimate the prediction of this epidemic in Italy.
We estimated the R0 parameter by means of a simple linear regression ad in https://kingaa.github.io/clim-dis/parest/parest.html

fit1 <- lm(log(totale_casi)~t,data=dat_csv)

```{r cars}
show(dat_csv)
fit1 <- lm(log(totale_casi)~t,data=dat_csv)
summary(fit1)
```

Estimates

You can also embed plots, for example:

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_casi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95%IC 
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario.
We fix a series of initial parameters.
```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-max(dat_csv$totale_casi); I0
# initial number of recovered
R0<-max(dat_csv$dimessi_guariti); R0
#beta 
beta0<-R_0/(14)
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r first scenario plot, echo=TRUE}
library(EpiDynamics)
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)

f2<-5
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)

f3<-3
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)

plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```