### 2020-03-14 draft_analysis_Italy.Rmd
Add home page, navigation and theme

---
title: "COVID19 - Forecast and predictions using a time dependent SEIR model" 
author: "PG"
date: "3/13/2020"
output: github_document
---
## The COVID dataset

The present analysis started from the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19, database provided by the Italian Civil Protection.  
```{r setup}
rm(list=ls())
###import italian dataset updated 13 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
days<-dim(dat_csv)[1]
dat_csv$t<-1:days
#database contains  the following variables
names(dat_csv)
# The total number of epidemic day is
days

```

With the aim to predict the future number of COVID19 cases on the basis of the actual data, we used a SEIR model applied to the COVID19 epidemic in Italy.
A first parameter to estimate is R0, the reproduction number (https://en.wikipedia.org/wiki/Basic_reproduction_number).   R0 indicates how contagious an infectious disease is. It is also referred to as "the reproduction number" of COVID19.

The R0 value has been estimated by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html.

The following plot presented the actual status.

```{r plots}
library(tidyr)
library(dplyr)
library(ggplot2)
df <- dat_csv %>%
  select(data, ricoverati_con_sintomi, terapia_intensiva, 
         totale_ospedalizzati, isolamento_domiciliare, 
         totale_attualmente_positivi, nuovi_attualmente_positivi, 
         dimessi_guariti, deceduti, totale_casi) %>%
  gather(key = "variable", value = "value", -data)
head(df, 3)

ggplot(df, aes(x = as.Date(data), y = value)) + 
  geom_line(aes(color = variable), size = 1) 
```

The plot shows an exponential grow of cases.\\

We estimate the R0 parameter by means of a linear model.

Y_t= a + beta * t +e_t

where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.

The slope coefficient is used to estimate R0 as in the following formula:

R0=beta*incubation period.

The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days. Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/. However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects.

In the calculation we considered an "incubation period" of 14 days for two reasons:  
1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 
2) 14 days is the worst scenario because in this period low impact COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).

We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).

```{r R0 trend}
#calculate r0 based in the last XX observation 
#vector for beta and standard deviation
beta_vec<-NULL
sd_vec<-NULL
#for cycle for all days to only the last 5.
for (i in 1:(days-4)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

label <-days:5

mean  <- (beta_vec*14)
lower <- ((beta_vec-1.96*sd_vec)*14)
upper <- ((beta_vec+1.96*sd_vec)*14)

df <- data.frame(label, mean, lower, upper)[order(label),]
df$label <- factor(df$label, levels=rev(df$label))

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() + 
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
  theme_bw() 
print(fp)

```

The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponential increase.  
However,  R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0. The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
```{r R0 forecast}
library(splines)
time<-1:(days-4)
# splines with only 3 basis, weighted for standard deviation of estimates.
beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)

r0.predict<-beta.predict
r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14
# plot the points (actual observations), regression line, and confidence interval
p <- ggplot(r0.predict, aes(date,beta_vec))
p <- p + geom_point() +labs(x="Date",y="R0 value") 
p <- p + geom_line(aes(date,fit))
p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
p
```

R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.

We want to make a short term forecast (14 days) with 3 scenario:

-Scenario 1: 10 exposed people for each COVID-19 case (no restrictions made or even no effects)

-Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)

-Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)

We made a forecast by means of a SEIR model fixing a series of initial status:  
- S0=N, Italian population   
- E= f * I0 (with f a fixed factor of the previous scenario)  
- I0: initial number of COVID-19 cases  
- R0: initial number of recovered

and parameters:  
- beta: the quantity connected to R0 is considered to vary according the previous estimation  
- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
- sigma0: the coronavirus transmission rate (half of flu epidemic)   
- mu0: the overall mortality rate

```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
# initial number of recovered
R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:

<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>

where the parameter beta here is time dependent, as estimated before by the gamma regression model.
```{r  scenario plot, echo=TRUE}
library(EpiDynamics)

# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
forecast<-14
seir1<-seir2<-seir3<-NULL
for(i in 1:forecast){
parameters <- c(mu = mu0, beta = beta.predict$fit[days+i], sigma = sigma0, gamma = 1/duration)
f1<-10
if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
f2<-5
if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
f3<-3
if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
}

date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
plot(date,c(dat_csv$totale_attualmente_positivi,seir1$I*N),type="l",ylab="Cases",xlab="time",main="Infected")
lines(date,c(dat_csv$totale_attualmente_positivi,seir2$I*N),col=2)
lines(date,c(dat_csv$totale_attualmente_positivi,seir3$I*N),col=3)
lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

The 3 scenarios show different numbers. 
If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$I*N)[length(seir2$I)]`).

In the next plot the cumulative number of infected.  
At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r ((seir2$I+seir3$R)*N)[length(seir2$I)]`).

```{r  cumulative plot, echo=TRUE}

plot(date,c(dat_csv$totale_casi,(seir1$I+seir1$R)*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
lines(date,c(dat_csv$totale_casi,(seir2$I+seir2$R)*N),col=2)
lines(date,c(dat_csv$totale_casi,(seir3$I+seir3$R)*N),col=3)
lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

```{r  forecast plot, echo=TRUE}

plot(date[(days+1):(days+forecast)],seir2$E*N,type="l",ylab="Cases",xlab="Date",main="14 days - Forecast Scenario 2",ylim=c(0,max(seir2$E*N,seir2$I*N,seir2$R*N )))
lines(date[(days+1):(days+forecast)],seir2$I*N,col=2)
lines(date[(days+1):(days+forecast)],seir2$R*N,col=3)
legend("topleft",c("Exposed","Infected","Recovered"),lty=1,col=1:4)
```