### 2020-03-15 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

```{r echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
knitr::opts_chunk$set(error = FALSE)
```
---
title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:html_document:
  df_print: paged
pdf_document: 
  toc: yes
---

<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

# Disclaimer

* We want to investigate the evolution of the coronavirus pandemic in Italy from a statistical perspective using aggregated data.

* Our point of view is that of surveillance with the goal of detecting important changes in the underlying (random) process as soon as possible after it has occured.

* We use data provided by  Italian Civil Protection Department

* This document is in a draft mode, and it is continuously updated.

* The layout of the draft must definitely be improved.

## The COVID dataset

The present analysis started from the dataset on COVID19 updated in
<https://github.com/pcm-dpc/COVID-19>, database provided by the Italian
Civil Protection.

# Software

Install  packages `dygraphs`,  `xts` and `EpiDynamics` if not available

```{r echo=TRUE}
checkpackage <- function(package) {
  if (!package %in% installed.packages()) install.packages(package)
}
checkpackage("dygraphs")
checkpackage("xts")
checkpackage("EpiDynamics")
checkpackage("webshot")
checkpackage("bsts")
```
and load them.
```{r, warning=FALSE}
library(dygraphs)
library(xts)
library(EpiDynamics)
library(webshot)
library(bsts)
```

# Source of the data

Download the data from

https://github.com/pcm-dpc/COVID-19/

# Results

## Load dataset
```{r}

rm(list=ls())
###import italian updated dataset 
  dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
days<-dim(dat_csv)[1]
dat_csv$t<-1:days
# The total number of epidemic day is
days

```
Several outcomes can be potentially monitored, that is

```{r}
names(dat_csv[,-c(1:2,13)])

```

It is worth noting that some outcomes present negative counts in some regions. It looks like some of these negative counts are redesignations. Outcomes presenting negative values cannot be analyzed using the proposed model.

Then we extract the timeseries.

```{r}
myDateTimeStr1 <- dat_csv$data
myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
days_dy<-as.Date(myPOSIXct1)
dat_csv_dy<-xts(dat_csv[,-c(1:2,12:13)], order.by = days_dy, frequency = 7)
```

```{r}
p <- dygraph(dat_csv_dy,main=paste("Italy",sep =""),xlab="Day",height=400,width=800) 
p
```

### The S(E)IR model (to be revised)

With the aim of predicting the future number of COVID19 cases on the
basis of the actual data, we used a SEIR model applied to the COVID19
epidemic in Italy.

We will consider  the classical [SIR model](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology) [@Kermack1927].

The model divides a population of hosts into three classes:
susceptible, infected, recovered.
The model describes how the portion of the population in each of these classes changes with time.
Births are modeled as flows from "elsewhere" into the susceptible class;
deaths are modeled as flows from the $S$, $I$, or $R$ compartment into "elsewhere".
If $S$, $I$, and $R$ refer to the numbers of individuals in each compartment, then these **state variables** change according to the following system of differential equations:
$$\begin{aligned}
\frac{d}{dt}S(t) &= B(t)-\lambda\,S(t)-\mu\,S(t)\\
\frac{d}{dt}I(t) &= \lambda\,S(t)-\gamma\,I(t)-\mu\,I(t)\\
\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t).\\
\end{aligned}$$
Here, $B$ is the crude birth rate (births per unit time), $\mu$ is the death rate and $\gamma$ is the recovery rate.
We'll assume that the force of infection, $\lambda$, for a constant population $N$
$$\lambda = \beta\,\frac{I}{N},$$
so that the risk of infection a susceptible faces is proportional to the *prevalence* (the fraction of the population that is infected).
This is known as the assumption of frequency-dependent transmission.

# The reproduction number of COVID19.

The number of infected individuals $I$ at time $t$ is approximately 
$$I(t)\;\approx\;I_0\,e^{R_0\,(\gamma+\mu)\,t}$$
where $I_0$ is the (small) number of infectives at time $0$, $\frac{1}{\gamma}$ is the infectious period, and $\frac{1}{\mu}$ is the host lifespan.

$R_0$ is  the reproduction
number (<https://en.wikipedia.org/wiki/Basic_reproduction_number>) and
indicates how contagious an infectious disease is.

Taking logs of both sides, we get

$$\log{I}(t)\;\approx\;\log{I_0}+(R_0)\,(\gamma+\mu)\,t,$$
which implies that a semi-log plot of $I$ vs $t$ should be approximately linear with a slope proportional to $R_0$ and the recovery rate.

```{r}
dat_csv_dy$log_totale_attualmente_positivi<-log(dat_csv_dy$totale_attualmente_positivi)
p <- dygraph(dat_csv_dy$log_totale_attualmente_positivi,main=paste("Italy"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
p
```

We estimate the $R_0$ parameter in the linear model.

$$
\log(I(t))= \alpha + \beta  t +e_t
$$

The estimated slope coefficient $\hat\beta$ is used to estimate $R_0$ as in the following formula:

$$\widehat\beta=(\widehat{R_0})\,(\gamma+\mu)$$
The parameter $\mu$<<$\gamma$ and it can not be considered. As consequence, R0 can be estimated as follows
$$\hat{R_0}=\frac{\hat{\beta}}{\gamma}
$$

The incubation period $1/ \gamma$ for the coronavirus is in mean 5.1 days with a range from 2-14 days. Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/. However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects.

In the calculation we considered an "incubation period" of 14 days for two reasons:  
1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 
2) 14 days is the worst scenario because in this period low impact COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).

We calculate several R0 values, each one based on a mobile window of 5 days, that can be sufficient to estimate a local trend, in order to assess if the R0 trend is decreasing (how is expected to be).

```{r R0 trend}
#calculate r0 based with a mobile window of 5 days
#vector for beta and standard deviation
beta_vec<-NULL
sd_vec<-NULL
#for cycle for R0 estimates from days-2 to days+2
for (i in 3:(days-2)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

label<-as.Date(substr(dat_csv$data,1,10))[3:(days-2)]

mean  <- (beta_vec*14)
lower <- ((beta_vec-1.96*sd_vec)*14)
upper <- ((beta_vec+1.96*sd_vec)*14)

df <- data.frame(label, mean, lower, upper)

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() +
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Date") + ylab("R0 Mean (95% CI)") +
  theme_bw() 
print(fp)

```

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-normal distribution for the beta (the slope) forcing its value to be greater than 0.
```{r R0 forecast}

time<-3:(days-2)
beta.model<-glm(log(beta_vec)~time,weights = 1/sd_vec,family=gaussian)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
beta.predict<-predict 
beta.predict[,3:5]<-exp(predict[,3:5])
r0.predict<-beta.predict
r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14
# plot the points (actual observations), regression line, and confidence interval
p <- ggplot(r0.predict, aes(date,beta_vec))
p <- p + geom_point() +labs(x="Date",y="R0 value") 
p <- p + geom_line(aes(date,fit))
p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
p
```

R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.

We want to make a short term forecast (14 days) with 3 scenario, based on the number of exposed people:

-Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)

-Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)

-Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)

We made a forecast by means of a SEIR model fixing a series of initial status:  
- S0=N, the size of Italian population   
- E= f * I0 (with f a fixed factor of the previous scenario)  
- I0: initial number of COVID-19 cases  
- R0: initial number of recovered

and parameters:  
- beta: the quantity connected to R0 is considered to vary according the previous estimation  
- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
- sigma0: the coronavirus transmission rate (half of flu epidemic)   
- mu0: the overall mortality rate

```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
# initial number of recovered
R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:

<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>

We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.   
We made a forecast forward of 14 days.

```{r bsts model fit and prediction, echo=TRUE}
# Bayesian Structural Time Series
ss <- AddLocalLinearTrend(list(), log(beta_vec))
model1 <- bsts(log(beta_vec),
               state.specification = ss,
               niter = 200)
par(mfrow = c(1,1))
plot(model1, "components", ylab="Log beta coefficient", xlab="days")
#previsioni
pred1 <- predict(model1, horizon = 16, burn = 100)
par(mfrow = c(1,1))
plot(pred1 , ylab="Log beta coefficient",main="Data and predictions")
# matrix of beta coefficients
coef<-(exp(pred1$distribution)[,3:16])
par(mfrow = c(1,1))
```
For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
We save the results and plot a credible interval at 90% (5-95% percentile).

```{r  scenario plot , echo=TRUE}
library(EpiDynamics)
seir1_sim<-seir2_sim<-seir3_sim<-NULL
for(s in 1:dim(coef)[1]){
  # average number of single connections of an infected person
  # less contacts, less probability of new infections
  # we keep constant the other parameters
  forecast<-14
  seir1<-seir2<-seir3<-NULL
  for(i in 1:forecast){
    parameters <- c(mu = mu0, beta = matrix(coef[s,i]), sigma = sigma0, gamma = 1/duration)
    f1<-10
    if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
    if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
    seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
    f2<-5
    if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
    if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
    seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
    f3<-3
    if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
    if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
    seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
  }
  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))
  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
  seir3_sim<-rbind(seir3_sim,cbind(rep(s,forecast),seir3))
}
seir1_sim[,2]<-seir2_sim[,2]<-seir3_sim[,2]<-rep(1:forecast,dim(coef)[1])
colnames(seir1_sim)[1]<-colnames(seir2_sim)[1]<-colnames(seir3_sim)[1]<-"sim"

### confidence limits
I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.05)
I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.95)

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1
mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario 1 Credible (confidence ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p

### confidence limits - scenario 2
I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.05)
I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.95)

mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario 2 Credible (confidence ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p

### confidence limits - scenario 3
I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.05)
I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.95)

mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario 3 Credible (confidence ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

The 3 scenarios show different numbers. 
If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).

In the next plot the cumulative number of infected.  
At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).