### 2020-03-17 BYM_model/INLA_def_html_prova.Rmd

---
title: "COVID19 - Forecast and predictions using a BYM model in Italy"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  df_print: paged
pdf_document: 
  toc: yes 
---
```{r echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
knitr::opts_chunk$set(error = FALSE)
```
<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

# Disclaimer

* We want to investigate the evolution of the coronavirus pandemic in  Italy

* Our point of view is that of surveillance with the goal of detecting important changes in the underlying (random) process as soon as possible after it has occured.

* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Lombardy Region

* This document is in a draft mode, and it is continuously updated.

* The layout of the draft must definitely be improved.

NB: to compile this file please download the Italian Shapefile from the repository  
and set the working directory here  
```{r echo=TRUE}
setwd("/Users/Paolo/My_Website/public/COVID19_SEIR_model")  
```
## The COVID dataset

The present analysis started from the dataset on COVID19 updated in
<https://github.com/pcm-dpc/COVID-19>, database provided by the Italian
Civil Protection.

# Software

Install  packages `dygraphs`,  `xts` and `EpiDynamics` if not available

```{r echo=TRUE}
checkpackage <- function(package) {
  if (!package %in% installed.packages()) install.packages(package)
}
checkpackage("dygraphs")
checkpackage("xts")
checkpackage("ggplot2")
checkpackage("spdep")
checkpackage("maptools")
checkpackage("INLA")
checkpackage("rgdal")
```
and load them.
```{r, warning=FALSE}
#Load libraries
library("dygraphs")
library("xts")
library("ggplot2")
library("spdep")
library("maptools")
library("INLA")
library("rgdal")
```
# Datasets loading
```{r}
db_istat<-read.csv("https://raw.githubusercontent.com/Paolin83/COVID19_SEIR_model/master/BYM_model/Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
###import international dataset
dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])

###import updated italian dataset  for each province
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
#select a Region, in this example "Lombardia" Code region 3
Region<-3
dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
dat_csv$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
#### number of province
nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])

#order the obsercation for date and space, useful for INLA later....
dat_csv<-dat_csv[order(dat_csv$data),]
dat_csv<-dat_csv[order(dat_csv$codice_provincia),]
###space ID
dat_csv$ID<-as.factor(dat_csv$codice_provincia)
levels(dat_csv$ID)<-1:length(levels(dat_csv$ID))
dat_csv$ID<-as.numeric(dat_csv$ID)
#calculate new_cases Y ~ Poi(mu)
dat_csv$cumulative_cases<-dat_csv$totale_casi
dat_csv$totale_casi[dat_csv$data!="2020-02-24 18:00:00"]<-unlist(tapply(dat_csv$totale_casi,dat_csv$ID,diff))
dat_csv$totale_casi[dat_csv$totale_casi<0]<-0
#import dataset shapefile Italian province and adjancies matrix
#### IMPORTANT, SET THE Work Directory of the shapefile 
nc.province <- readShapePoly("BYM_model/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp")
nc.province<-nc.province[nc.province$COD_REG==Region,]
nc.province<-nc.province[order(nc.province$COD_PROV),]
nc.province$ID<-as.factor(nc.province$COD_PROV)
levels(nc.province$ID)<-1:length(levels(nc.province$ID))
nc.province$ID<-as.numeric(nc.province$ID)
#Create adjacency matrix
nc.nb <- poly2nb(nc.province,snap=0.01)
nb2INLA("nc.adj", nc.nb)
file.adj <- "nc.adj"
################################
```

# Spatial analysis with R-INLA and BYM 
We modelled COVID 19 cases by means of a BYM (Besag, York and Molli\'e) model using an Integrated Nested Laplace Approximation (INLA).  
* Starting phase

For the $i$-th Nuts-3 Region (Italian province), the observed number of COVID19, $y_{i}$, was modelled as follows:

$$
  Y_i|\lambda_i\sim \mbox{Poisson}(\lambda_i), \qquad i=1,2,\ldots \\
$$
with $i=1,\ldots,T$ and $\lambda_i>0$.
We modelled the incidence of COVID-19 by means of a BYM model including the population size as offset variable as:  
$$
	log(\frac{\lambda_{i}}{N_{i}})=\alpha+\mu_i+\nu_i, \\
$$
where $\alpha$ is the intercept, $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR) and $N_i$ is the population size of each NUTS-3 Region.