### 2020-03-19 BYM_model/INLA_Veneto_html.Rmd
Add home page, navigation and theme

---
title: "COVID19 - Forecast and predictions using a BYM model in Italy"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:html_document:  
  df_print: paged  
pdf_document:   
  toc: yes

always_allow_html: yes
---
```{r echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
knitr::opts_chunk$set(error = FALSE)
```
<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

# Disclaimer

* We want to investigate the evolution of the coronavirus pandemic in  Italy

* Our point of view is that of surveillance with the goal of detecting important changes in the underlying (random) process as soon as possible after it has occured.

* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Lombardy Region

* This document is in a draft mode, and it is continuously updated.

* The layout of the draft must definitely be improved.

NB: to compile this file please download the Italian Shapefile from the repository  
and set the working directory here  
```{r echo=TRUE}
setwd("/Users/Paolo/My_Website/public/COVID19_SEIR_model")  
```
## The COVID dataset

The present analysis started from the dataset on COVID19 updated in
<https://github.com/pcm-dpc/COVID-19>, database provided by the Italian
Civil Protection.

# Software

Install  packages `dygraphs`,  `xts` and `EpiDynamics` if not available

```{r echo=TRUE}
checkpackage <- function(package) {
  if (!package %in% installed.packages()) install.packages(package)
}
checkpackage("dygraphs")
checkpackage("xts")
checkpackage("ggplot2")
checkpackage("spdep")
checkpackage("maptools")
checkpackage("INLA")
checkpackage("rgdal")
```
and load them.
```{r, warning=FALSE}
#Load libraries
library("dygraphs")
library("xts")
library("ggplot2")
library("spdep")
library("maptools")
library("INLA")
library("rgdal")
```
# Datasets loading
```{r}
db_istat<-read.csv("https://raw.githubusercontent.com/Paolin83/COVID19_SEIR_model/master/BYM_model/Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
###import international dataset
dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])

###import updated italian dataset  for each province
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
#select a Region, in this example "Veneto" Code region 3
Region<-5
dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
dat_csv$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
#### number of province
nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])

#order the obsercation for date and space, useful for INLA later....
dat_csv<-dat_csv[order(dat_csv$data),]
dat_csv<-dat_csv[order(dat_csv$codice_provincia),]
###space ID
dat_csv$ID<-as.factor(dat_csv$codice_provincia)
levels(dat_csv$ID)<-1:length(levels(dat_csv$ID))
dat_csv$ID<-as.numeric(dat_csv$ID)
#calculate new_cases Y ~ Poi(mu)
dat_csv$cumulative_cases<-dat_csv$totale_casi
dat_csv$totale_casi[dat_csv$data!="2020-02-24 18:00:00"]<-unlist(tapply(dat_csv$totale_casi,dat_csv$ID,diff))
dat_csv$totale_casi[dat_csv$totale_casi<0]<-0
#import dataset shapefile Italian province and adjancies matrix
#### IMPORTANT, SET THE Work Directory of the shapefile 
nc.province <- readShapePoly("BYM_model/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp")
nc.province<-nc.province[nc.province$COD_REG==Region,]
nc.province<-nc.province[order(nc.province$COD_PROV),]
nc.province$ID<-as.factor(nc.province$COD_PROV)
levels(nc.province$ID)<-1:length(levels(nc.province$ID))
nc.province$ID<-as.numeric(nc.province$ID)
#Create adjacency matrix
nc.nb <- poly2nb(nc.province,snap=0.01)
nb2INLA("nc.adj", nc.nb)
file.adj <- "nc.adj"
################################
```

# Spatial analysis with R-INLA and BYM 
We modelled COVID 19 cases by means of a BYM (Besag, York and Molli\'e) model using an Integrated Nested Laplace Approximation (INLA).  
* Starting phase

For the $i$-th Nuts-3 Region (Italian province), the observed number of COVID19, $y_{i}$, was modelled as follows:

$$
  Y_i|\lambda_i\sim \mbox{Poisson}(\lambda_i), \qquad i=1,2,\ldots \\
$$

and $\lambda_i>0$. We modelled the incidence of COVID-19 by means of a BYM model including the population size as offset variable as:

$$
 log(\frac{\lambda_{i}}{N_{i}})=\alpha+\mu_i+\nu_i, \\
$$
where $\alpha$ is the intercept, $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR) and $N_i$ is the population size of each NUTS-3 Region.

The parameter were estimated by INLA (Integrated Nested Laplace Approximation) and R software.

```{r}
nomi_prov<-levels(dat_csv$denominazione_provincia)
barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum),names =1:nprov,args.legend = list(x = "topright"),legend.text =paste(1:nprov,nomi_prov),ylab="Number fo cases")
```
The number of observed cases $y_i$ is very different among  NUTS-3 regions.

```{r}
pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:nprov,args.legend = list(x = "topright"),legend.text =paste(1:nprov,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
```
A correct comparison is the made by incidence cases $y_i/N_i$, here riported for x1000 inhabitants. 
#Spatial model 
We estimate a BYM model fixing prior distributions for iid and besag components as follows:
```{r}
prior.iid = c(1,0.01)
prior.besag = c(1,0.001)
initial.iid = 4
initial.besag = 3
```
We create new variables that are required by INLA procedure. 
```{r}
##########################indexing, for 
#### number of days
dat_csv$t<-as.numeric(dat_csv$data)
days<-max(dat_csv$t)
dat_csv$t3<-dat_csv$t2<-dat_csv$t
#### province
dat_csv$ID2<-dat_csv$ID
########################## formula BYM
formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
fit_1<-inla(formula.bym, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_1)
##incidence rate ratio for each province
IRR_mean <- exp(fit_1$summary.random$ID$mean[1:nprov])
nc.province$IRR_mean<-IRR_mean
spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios")
nomi_prov
```
The image reported the IRR=exp($\mu_i$), the increase of the Incidence of Covid-19 in each NUTS-3 Region respect to the overall mean.

#Temporal model 
We consider the temporal aspect estimating a Bayesian RW2 model to model the temporal trend of COVID-19 in the reported temporal window. The observed number of COVID-19 a the time $t$, $y_{t}$, is modelled as follows  
$$
  Y_t|\lambda_t\sim \mbox{Poisson}(\lambda_t), \qquad t=i,2,\ldots \\
$$

with t=1,...,T, the quantity $\lambda_{t}$, is modelled by a Random Walk of order 2 (RW2) variable  
$$
  log{\lambda_{it}}=\alpha+\gamma_t+\phi_t,  \\
$$

where $\alpha$ is the intercept, $\gamma_t \sim RW2$  are the coefficients related to the random walk process while $\phi_t \sim N(0, \tau_t^{-1})$ are the temporal specific random errors.

```{r}
Date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days-1,1)
plot(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum),xlab="Date",ylab="New cases",type="l")
points(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum))
formula_t = totale_casi ~ 1+f(t,model="rw2", constr = FALSE)+f(t2,model="iid")
fit_2<-inla(formula_t, family="poisson", data=dat_csv,control.compute = list(dic=T))
summary(fit_2)
##Incidence rate ratios, increase od cases respect to time 1
plot(Date,exp(fit_2$summary.random$t$mean),ylab="temporal IRR",xlab="Date")

label<-Date

mean  <- exp(fit_2$summary.random$t$mean[1:days]+fit_2$summary.fixed$mean)*nprov
lower <- exp(fit_2$summary.random$t$mean[1:days]-1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov
upper <- exp(fit_2$summary.random$t$mean[1:days]+1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov

df <- data.frame(label, mean, lower, upper)

fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() +
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Date") + ylab("Estimated new cases") +
  theme_bw() 
print(fp)

```
We now consider a BYM model with spatial and temporal components and their complete interactions (Type IV). The observed number of COVID-19, $y_{it}$, was modelled as:  
$$
  Y_{it} \sim Poisson (\lambda_{it}),  \\
$$

with i=1,...,`r nprov`and t=1,...,`r days`.  
We modelled the ratio $\frac{\lambda_{it}}{N_{i}}$, the Incidence of COVID-19, with a BYM model including temporal covariates as follows
$$
  log(\frac{\lambda_{it}}{N_{i}})=\alpha+\mu_i+\nu_i+\gamma_t+\phi_t+\delta_{it},  \\
$$
  where $\alpha$ is the intercept and:

*	space: $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR);

*	time:$\gamma_t \sim RW2$  are the coefficients related to the random walk process, while $\phi_t \sim N(0, \tau_t^{-1} )$ are the temporal specific random errors.

*	space-time: $\delta_{it}$ are the coefficients related space-time interactions that in the its type 4 formulation (please see Spatial and Spatio-temporal Bayesian Models with R - INLA, Blangiardo and Cameletti) is made by the Kronecker product of time and space indexes.

```{r}
####specification
formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
  f(t,model="rw2",constr = FALSE) +
  f(t2,model="iid") +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))
#estimation
fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_st4)
```

The time and space component mean estimates can be extracted here

```{r}
#overall trend
t_1<-fit_st4$summary.random$t$mean 
#iid trend
t_2<-fit_st4$summary.random$t2$mean 
 # spatial iCAR component
s_1<-fit_st4$summary.random$ID$mean[1:nprov]
# spatial idd
s_2<-fit_st4$summary.random$ID$mean[(nprov+1):(nprov*2)] 
#spatio-temporal model with car and ar1 specification
st<-fit_st4$summary.random$ID2$mean 
```

When can represent the IRR(=exp($\mu_i+\frac{1}{T} \sum_{i=t}^{T} \delta_{it}$)) for each NUTS-3 regions.

```{r}
nc.province$IRR_st<-exp(s_1+apply(t(matrix(st,nrow=nprov)),2,mean))
l1 = list("sp.text", coordinates(nc.province), nc.province$DEN_PROV,col="grey", cex=0.7,font=2)
spplot(nc.province, c( "IRR_st"),main="IRR", sp.layout=list(l1))
```
  Trend for each province is reported below.

```{r}
##overall trend
plot(Date,exp(t_1),ylab="IRR",main="Overall time trend")

## time trend for each province
db_plot<-data.frame(t_1+t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Venezia"
province_trend<-xts(db_plot, order.by = Date, frequency = 7)
p <- dygraph(province_trend,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
p

#space time variation for each province from overall trend
db_plot<-data.frame(t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Venezia"
province_trend<-xts(db_plot, order.by = Date, frequency = 7)
p <- dygraph(province_trend,main=paste("Veneto Region Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
p
```

#Predictions 3 days forward of new COVID-19 cases
We made a forecast of 3 days forward on the number of COVID19 by means of BYM model and considering as covariate the number of Hubei epidemic lagged of 8 days because the different moment of restrictive measures between Italy and Hubei.
```{r}
Forecast=3
dat_csv2<-dat_csv 
dat_csv2$t<-rep(1:days+days,nprov)
dat_csv2$totale_casi<-NA
dat_csv2<-dat_csv2[dat_csv2$t<=(days+Forecast),]
dat_csv_n<-rbind(dat_csv,dat_csv2)
dat_csv_n$t3<-dat_csv_n$t2<-dat_csv_n$t
#introducting hubei starting on 22/01, quarantine started on 25/01
# applying a smoothing , data has errors
hubei_new_cases<-filter(diff(hubei), rep(1, 5))

plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
abline(v=2,col=1,lty=2)
abline(v=4,col=2,lty=2)
abline(v=10,col=3,lty=3)
legend("bottomright",c("Hubei","Veneto Region"),lty=1,col=1:2)
text(10,1500,"Hubei quarantine")
text(12,1350,"Lodi home lock",col=2)
text(20,1200,"Veneto home lock",col=3)
#province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
#other provinces on 11/03 lagged started their restrictions
hubei_lodi<-hubei_new_cases[1:(days+Forecast)]
hubei_others<-c(rep(NA,8),hubei_new_cases[1:(days+Forecast-8)])
dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
dat_csv_n$hubei<-rep(hubei_others,nprov)
#overdispersion
dat_csv_n$over<-1:dim(dat_csv_n)[1]
```
We model the number of cases.

In the rw2 I set constr = FALSE is set to FALSE and that, for this reason, the intercept is not included in the linear predictor. 
```{r}
## include overdispersion parameter
formula.intIVn<- totale_casi ~ log(hubei+1)+log(pop)+
  f(t,model="rw2", constr = FALSE) +
  f(t2,model="iid") +f(over,model="iid") +
f(ID,model="bym",graph=file.adj) +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))+ f(over,model="iid")

fit_st4n<-inla(formula.intIVn, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
summary(fit_st4n)
```

```{r}
#### components
#overall trend
t_1<-fit_st4n$summary.random$t$mean 
#iid trend
t_2<-fit_st4n$summary.random$t2$mean 
 # spatial CAR componen
s_1<-fit_st4n$summary.random$ID$mean[1:nprov]
# spatial idd
s_2<-fit_st4n$summary.random$ID$mean[(nprov+1):(nprov*2)] 
# spatio-temporal coefficients 
st<-fit_st4n$summary.random$ID2$mean 
## Trends for province
Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
## time trend for each province
db_plot<-data.frame(t_1+t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Milano"
province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
p <- dygraph(province_trend,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
p

#space time variation for each province from overall trend
db_plot<-data.frame(t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Venezia"
province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
p <- dygraph(province_trend,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
p
#extract number of cases for Verona

#### FOR Verona as for example
est_verona<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Verona",]

days.before<-Date_n[1:days]
days.ahead<-Date_n[(days+1):(days+Forecast)]
mu.lower<-est_verona$mean-1.96*est_verona$sd
mu.lower[mu.lower<0]<-0
mu.upper<-est_verona$mean+1.96*est_verona$sd
mu.med<-xts(est_verona$mean,order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
step.ahead<-Forecast
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Verona (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```